---
title: "evaporation"
author: "Kevin Li"
date: "10/26/2021"
output: html_document
---

```{r setup, message= FALSE}

library(tidyverse)
library(lubridate)
library(egg)
library(mgcv)
library(PerformanceAnalytics) # For chart.Correlation
library(DHARMa)

```

## Evaporation data from raw

Just exploring evaporation starting from raw data. First combine the control and plot data into one unified dataset with time and date. 

```{r combine_raw}
evap <- read.csv("./data/evap_raw.csv")
evap_ctrl <- read.csv("./data/evap_control.csv")
rain <- readRDS("Data/dailyrain.rds")

evap_all <- evap %>% select(date:evpRate_g.mm2.s) %>%
  filter(paperID != "") %>%
  bind_rows(
    evap_ctrl %>% select(date:evpRate_g.mm2.s) %>%
      filter(startTime != "") %>%
      rename(paperID = ID)
  ) %>%
  mutate(startTime = as.POSIXlt(paste(date, startTime, sep = ":"), 
                             format = "%m/%d/%Y:%H:%M:%S")) %>% # I convert the time and date columns to r formats
  mutate(endTime = as.POSIXlt(paste(date, endTime, sep = ":"), 
                             format = "%m/%d/%Y:%H:%M:%S")) %>%
  mutate(date = as.Date(date, "%m/%d/%Y")) %>%
  rowwise() %>% 
  mutate(dateTime = mean(c(startTime, endTime))) %>% # the date/time at the center of the drying period
  ungroup() %>%
  mutate(decTime = hour(dateTime) + 
           minute(dateTime)/60 + 
           second(dateTime)/3600) %>% # the decimal time at the center of the drying period
  mutate(hr.2 = round(decTime/2,0)) %>%
  mutate(evapRate_ng = ifelse(evpRate_g.mm2.s > 0, evpRate_g.mm2.s*1e9, 1)) %>%  # evaporation rate in nanograms, so the units aren't so small
  mutate(locID = factor(ifelse(is.na(plotID), type, plotID))) %>%  # locID and locType are just to help with plotting
  mutate(locType = ifelse(is.na(plotID), type, "plot")) %>%
  mutate(dateFac = factor(date)) %>%
  mutate(dateNum = as.numeric(date - as.Date("05/30/2016", "%m/%d/%Y"))) %>%
  left_join(rain %>% 
              mutate(rain_lag = dplyr::lag(rain_mm, order_by = date)) %>%
              select(date, rain_mm, rain_lag))
```

## Plot some of the data

```{r loess_date, message=FALSE}
ggplot(evap_all, 
       aes(date, evapRate_ng, color = plotID, shape = locType, group = locType)) + 
  geom_point() + geom_smooth() + 
  scale_shape_manual(values = c(0:2, 19)) + 
  scale_color_distiller(palette = "Spectral") + 
  facet_wrap(~ locType)

```

Here I plot the three control datasets separately and all the plots together against the dates. I fit a LOESS smoothing to the data, which is like a moving window type of analysis, as I understand it. These are the raw data points, not averages.

### Mean evaporation rate by date

```{r spaghetti_mn}
ggplot(evap_all %>% group_by(date, locID, plotID, locType) %>% 
         summarize(evap_mn = mean(evapRate_ng)), 
       aes(date, evap_mn, color = plotID, shape = locType, group = locID)) + 
  geom_line() + geom_point() +
  scale_shape_manual(values = c(0:2, 19)) + 
  scale_color_distiller(palette = "Spectral") + 
  facet_wrap(~ locType)
```

I've taken the mean of the evaporation rate on each date, by plot or control location. Individual plots or control points are connected by a line.

### Time of day

```{r gam_time}
ggplot(evap_all, 
       aes(decTime, evapRate_ng, color = plotID, shape = locType, group = locType)) + 
  geom_point() + geom_smooth(method = "gam") + 
  scale_shape_manual(values = c(0:2, 19)) + 
  scale_color_distiller(palette = "Spectral") + 
  facet_wrap(~ locType)
```

I've fit a GAM to the raw (unaveraged) data over time of day over the control locations and the plots (taken together)

### Time of day, over months

```{r gam_time_month}

ggplot(evap_all %>% mutate(month = month(date)), 
       aes(decTime, evapRate_ng, color = locType, shape = locType)) + 
  geom_point() + geom_smooth(method = "gam") + 
  scale_shape_manual(values = c(0:2, 19)) + 
  facet_wrap(~ month) + theme_article()

```

I've fit a GAM over time for control locations or plots, but separately by months, to see if the nonlinear relationship changes over time.

## Model

Linear mixed models

```{r lmer_models}
library(lme4)
library(lmerTest)

evap_lmer <- lmer(log(evapRate_ng) ~ poly(decTime, 2) + (1|dateFac) + (1|locID),
          data = evap_all, 
          REML = F)

evap_lmer_polyre1 <- lmer(log(evapRate_ng) ~ poly(decTime, 2) + (poly(decTime, 1)|dateFac) + (1|locID),
          data = evap_all, 
          REML = F)

evap_lmer_polyre2 <- lmer(log(evapRate_ng) ~ poly(decTime, 2) + (poly(decTime, 2)|dateFac) + (1|locID),
          data = evap_all, 
          REML = F)

evap_lmer_polyre1v2 <- lmer(log(evapRate_ng) ~ poly(decTime, 2) + (poly(decTime, 1)|dateFac) + (poly(decTime, 1)|locID),
          data = evap_all, 
          REML = F)

evap_lmer_polyre1v3 <- lmer(log(evapRate_ng) ~ poly(decTime, 2) + (1|dateFac) + (poly(decTime, 1)|locID),
          data = evap_all, 
          REML = F)


performance::compare_performance(evap_lmer, evap_lmer_polyre1, evap_lmer_polyre2, evap_lmer_polyre1v2, evap_lmer_polyre1v3)

testlmer <- simulateResiduals(evap_lmer_polyre1, n=1000)

testlmer.grp <- recalculateResiduals(testlmer, group = evap_all$locID)


evap_lmer_datenum <- lmer(log(evapRate_ng) ~ poly(decTime, 2) + dateNum + (1|locID),
          data = evap_all, 
          REML = F)

evap_lmer_datenum_poly <- lmer(log(evapRate_ng) ~ poly(decTime, 2) * dateNum + (1|locID),
          data = evap_all, 
          REML = F)

# evap_glmer <- glmer(evapRate_ng ~ poly(decTime, 2) + (1|dateFac) + (1|locID),
#           data = evap_all, family = Gamma(link = "identity"))
# 
# evap_glmer_polyre <- glmer(evapRate_ng ~ poly(decTime, 2) + (poly(decTime, 2)|dateFac) + (1|locID),
#           data = evap_all, family = Gamma(link = "inverse"))
# 
# test_evap_lmer_polyre <- simulateResiduals(evap_lmer_polyre)

```


View prediction curves by plots. Curves show estimated prediction based on the "middle" date for the data points taken for that plot

```{r predictions_plot}

dateloccombo <- evap_all %>% 
  group_by(locID) %>% 
  summarize(dateFac = factor(quantile(date, p = 0.5, type = 1)), .groups = "drop")

newdata_dateloc <- dateloccombo %>% 
  slice(rep(1:n(), times = length(seq(6,15,by = 0.5)))) %>%
  mutate(decTime = rep(seq(6,15,by = 0.5), each = nrow(dateloccombo)))

pred_evap_0 <- predict(evap_lmer, newdata = newdata_dateloc)
pred_evap_p1 <- predict(evap_lmer_polyre1, newdata = newdata_dateloc)
pred_evap_p2 <- predict(evap_lmer_polyre2, newdata = newdata_dateloc)
pred_evap_p1v2 <- predict(evap_lmer_polyre1v2, newdata = newdata_dateloc)
pred_evap_p1v3 <- predict(evap_lmer_polyre1v3, newdata = newdata_dateloc)
newdata_dateloc$pred.evap.0 = exp(pred_evap_0)
newdata_dateloc$pred.evap.p1 = exp(pred_evap_p1)
newdata_dateloc$pred.evap.p2 = exp(pred_evap_p2)
newdata_dateloc$pred.evap.p1v2 = exp(pred_evap_p1v2)
newdata_dateloc$pred.evap.p1v3 = exp(pred_evap_p1v3)

ggplot(newdata_dateloc, aes(decTime)) + 
  geom_line(aes(y = pred.evap.0), color = "grey", alpha = 0.4) + 
  geom_line(aes(y = pred.evap.p1), color = "blue", alpha = 0.4) + 
  geom_line(aes(y = pred.evap.p2), color = "red", alpha = 0.4) +
  geom_line(aes(y = pred.evap.p1v2), color = "green", alpha = 0.4) +
  geom_line(aes(y = pred.evap.p1v2), color = "purple", alpha = 0.4) +
  geom_point(data = evap_all, mapping = aes(decTime, evapRate_ng), inherit.aes = F) + 
  facet_wrap(~locID) + theme_article()

```


```{r predictions_date}

datelocunq <- evap_all %>% 
  group_by(dateFac, locID) %>% 
  summarize(n = n(), .groups = "drop")

newdata_date <- datelocunq %>% 
  select(dateFac, locID) %>% 
  slice(rep(1:n(), times = length(seq(6,15,by = 0.5)))) %>%
  mutate(decTime = rep(seq(6,15,by = 0.5), each = nrow(datelocunq)))

dpred_evap_0 <- predict(evap_lmer, newdata = newdata_date)
dpred_evap_p1 <- predict(evap_lmer_polyre1, newdata = newdata_date)
dpred_evap_p2 <- predict(evap_lmer_polyre2, newdata = newdata_date)
newdata_date$pred.evap.0 = exp(dpred_evap_0)
newdata_date$pred.evap.p1 = exp(dpred_evap_p1)
newdata_date$pred.evap.p2 = exp(dpred_evap_p2)

newdata_date <- newdata_date %>% left_join(evap_all %>% select(locID, locType) %>% distinct())

ggplot(newdata_date, aes(decTime, group = locID, color = locType)) + 
  # geom_line(aes(y = pred.evap.0), color = "grey", alpha = 0.2) + 
  # geom_line(aes(y = pred.evap.p1), color = "blue", alpha = 0.2) + 
  # geom_line(aes(y = pred.evap.p2), color = "red", alpha = 0.2) +
  geom_line(aes(y = pred.evap.p1), alpha = 0.5) + 
  geom_point(data = evap_all, mapping = aes(decTime, evapRate_ng, color = locType), inherit.aes = F) + 
  facet_wrap(~dateFac) + theme_article()

ggplot(newdata_date %>% filter(locType == "plot"), aes(decTime, color = locID)) + 
  # geom_line(aes(y = pred.evap.0), color = "grey", alpha = 0.2) + 
  # geom_line(aes(y = pred.evap.p1), color = "blue", alpha = 0.2) + 
  # geom_line(aes(y = pred.evap.p2), color = "red", alpha = 0.2) +
  geom_line(aes(y = pred.evap.p1), alpha = 0.8) + 
  geom_point(data = evap_all %>% filter(locType == "plot"),
             mapping = aes(decTime, evapRate_ng, color = locID), inherit.aes = F) + 
  facet_wrap(~dateFac) + theme_article() + theme(legend.position = "none")

ggplot(newdata_date %>% filter(locType == "control1"), aes(decTime, group = locID)) + 
  # geom_line(aes(y = pred.evap.0), color = "grey", alpha = 0.2) + 
  # geom_line(aes(y = pred.evap.p1), color = "blue", alpha = 0.2) + 
  # geom_line(aes(y = pred.evap.p2), color = "red", alpha = 0.2) +
  geom_line(aes(y = pred.evap.p1), alpha = 0.8) + 
  geom_point(data = evap_all %>% filter(locType == "control1"),
             mapping = aes(decTime, evapRate_ng), inherit.aes = F) + 
  facet_wrap(~dateFac) + theme_article() + theme(legend.position = "none")

ggplot(newdata_date %>% filter(locType == "gazebo"), aes(decTime, group = locID)) + 
  # geom_line(aes(y = pred.evap.0), color = "grey", alpha = 0.2) + 
  # geom_line(aes(y = pred.evap.p1), color = "blue", alpha = 0.2) + 
  # geom_line(aes(y = pred.evap.p2), color = "red", alpha = 0.2) +
  geom_line(aes(y = pred.evap.p1), alpha = 0.8) + 
  geom_point(data = evap_all %>% filter(locType == "gazebo"),
             mapping = aes(decTime, evapRate_ng), inherit.aes = F) + 
  facet_wrap(~dateFac) + theme_article() + theme(legend.position = "none")

ggplot(newdata_date %>% filter(locType == "house"), aes(decTime, group = locID)) + 
  # geom_line(aes(y = pred.evap.0), color = "grey", alpha = 0.2) + 
  # geom_line(aes(y = pred.evap.p1), color = "blue", alpha = 0.2) + 
  # geom_line(aes(y = pred.evap.p2), color = "red", alpha = 0.2) +
  geom_line(aes(y = pred.evap.p1), alpha = 0.8) + 
  geom_point(data = evap_all %>% filter(locID == "house"),
             mapping = aes(decTime, evapRate_ng), inherit.aes = F) +
  facet_wrap(~dateFac) + theme_article() + theme(legend.position = "none")

```

```{r predicted_values}
# based on model evap_lmer_polyre1

pred_dat <- data.frame(locID = rep(unique(evap_all$locID), each = length(unique(evap_all$dateFac))), 
                       decTime = 10, 
                       dateFac = rep(unique(evap_all$dateFac), times = length(unique(evap_all$locID))))

pred_dat$pred.evap.0 = exp(predict(evap_lmer, newdata = pred_dat))
pred_dat$pred.evap.1 = exp(predict(evap_lmer_polyre1, newdata = pred_dat))

# prediction at the population mean of day effects (at 10 AM)
pred_10AM <- data.frame(locID = unique(evap_all$locID), 
                       decTime = 10, 
                       dateFac = NA)

pred_10AM$pred.evap.1 = exp(predict(evap_lmer_polyre1, newdata = pred_10AM, 
                                    re.form = ~ (1|locID), allow.new.levels = TRUE))

ggplot(pred_dat) + 
  # geom_histogram(aes(x = pred.evap.0), color = "green", alpha = 0.2) + 
  geom_histogram(aes(x = pred.evap.1), color = "blue", alpha = 0.2) +
  geom_vline(data = pred_10AM, mapping = aes(xintercept = pred.evap.1)) +
  facet_wrap(~locID) + theme_article()

```

# model uncertainty based on data subsets
```{r uncertainty}

# randomly draw 90% data
set.seed(5)
n.rep = 1000

crossvaldat <- list()
crosspred <- list()
crossmods <- list()

for(i in 1:n.rep){
  sub.index <- sample(x = 1:nrow(evap_all), size = round(0.9*nrow(evap_all)))
  
  data.i <- evap_all[sub.index,]
  data.out <- evap_all[!1:nrow(evap_all) %in% sub.index, ]
  
  model.i0 <- lmer(log(evapRate_ng) ~ poly(decTime, 2) + (1|dateFac) + (1|locID), 
                   data = data.i, REML = F)
  model.i1 <- lmer(log(evapRate_ng) ~ poly(decTime, 2) + (poly(decTime, 1)|dateFac) +
                     (1|locID), data = data.i, REML = F)
  model.i2 <- lmer(log(evapRate_ng) ~ poly(decTime, 2) + (poly(decTime, 2)|dateFac) +
                     (1|locID), data = data.i, REML = F)
  
  new.i <- data.frame(locID = unique(data.i$locID), 
                       decTime = 10, 
                       dateFac = NA)
  
  new.i$pred0 = exp(predict(model.i0, newdata = new.i, re.form = ~ (1|locID), 
                            allow.new.levels = TRUE))
  new.i$pred1 = exp(predict(model.i1, newdata = new.i, re.form = ~ (1|locID), 
                            allow.new.levels = TRUE))
  new.i$pred2 = exp(predict(model.i2, newdata = new.i, re.form = ~ (1|locID), 
                            allow.new.levels = TRUE))
  
  predict.i0 <- exp(predict(model.i0))
  predict.i1 <- exp(predict(model.i1))
  predict.i2 <- exp(predict(model.i2))
  
  out.predict0 <- exp(predict(model.i0, newdata = data.out, allow.new.levels = TRUE))
  out.predict1 <- exp(predict(model.i1, newdata = data.out, allow.new.levels = TRUE))
  out.predict2 <- exp(predict(model.i2, newdata = data.out, allow.new.levels = TRUE))
  
  crossvaldat[[i]] <- bind_rows(data.i %>% select(locID, decTime, dateFac, evapRate_ng) %>% 
                                  mutate(cross.class = "train") %>%
                                  bind_cols(predict0 = predict.i0,
                                            predict1 = predict.i1,
                                            predict2 = predict.i2),
                                data.out %>% select(locID, decTime, dateFac, evapRate_ng) %>% 
                                  mutate(cross.class = "out") %>%
                                  bind_cols(predict0 = out.predict0,
                                            predict1 = out.predict1,
                                            predict2 = out.predict2)
                                ) %>% mutate(i = i)
  
  crosspred[[i]] <- new.i
  
  crossmods[[i]] <- list(model.i0, model.i1, model.i2)
}

pred_sum <- bind_rows(crosspred) %>% group_by(locID) %>%
  summarize(mn.pred0 = mean(pred0, na.rm = T), 
            sd.pred0 = sd(pred0, na.rm = T),
            mn.pred1 = mean(pred1, na.rm = T),
            sd.pred1 = sd(pred1, na.rm = T),
            mn.pred2 = mean(pred2, na.rm = T), 
            sd.pred2 = sd(pred2, na.rm = T),
            .groups = "drop") %>%
  pivot_longer(cols = mn.pred0:sd.pred2, names_to = c(".value", "model"), names_sep = "\\.") %>%
  left_join(evap_all %>% select(locID, locType) %>% distinct())

ggplot(pred_sum, aes(locID, mn, color = model)) +
  geom_pointrange(aes(ymin = mn - sd, ymax = mn + sd)) + 
  theme_article() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

rmse.fullmodel0 <- sqrt(mean(residuals(evap_lmer)^2))
rmse.fullmodel1 <- sqrt(mean(residuals(evap_lmer_polyre1)^2))
rmse.fullmodel2 <- sqrt(mean(residuals(evap_lmer_polyre2)^2))

# error characteristics
mcpred <- bind_rows(crossvaldat)

rmse.out <- mcpred %>% 
  mutate(error0 = predict0 - evapRate_ng,
         error1 = predict1 - evapRate_ng,
         error2 = predict2 - evapRate_ng) %>%
  group_by(cross.class, i) %>% 
  summarize(rmse0 = sqrt(mean(error0^2)),
            rmse1 = sqrt(mean(error1^2)),
            rmse2 = sqrt(mean(error2^2)),
            .groups = "drop") %>%
  pivot_wider(id_cols = i, names_from = cross.class, values_from = rmse0:rmse2) %>%
  mutate(diff0 = rmse0_out - rmse0_train,
         diff1 = rmse1_out - rmse1_train,
         diff2 = rmse2_out - rmse2_train)

rmse.long <- rmse.out %>% select(-(diff0:diff2)) %>%
  pivot_longer(cols = rmse0_out:rmse2_train,
               names_to = c("model", ".value"), names_sep = "_")

ggplot(rmse.long) +
  geom_histogram(aes(x = out), fill = "orange", alpha = 0.2) +
  geom_histogram(aes(x = train), fill = "dark green", alpha = 0.2) + 
  facet_wrap(~model) + theme_article()

diff.long <- rmse.out %>% select(i, diff0:diff2) %>%
  pivot_longer(cols = diff0:diff2, names_to = "model", values_to = "diff")

ggplot(diff.long, aes(x = diff)) + geom_histogram() + 
  geom_vline(xintercept = 0, color = "red") + facet_wrap(~model)

ggplot(diff.long %>% group_by(model) %>% summarize(mean = mean(diff), 
                                                   CI2.5 = quantile(diff, 0.0275),
                                                   CI97.5 = quantile(diff, 0.975))) +
  geom_vline(xintercept = 0) +
  geom_pointrange(aes(x = mean, y = model, xmin = CI2.5, xmax = CI97.5))

```

# Look at the relationship between standardized evap and other abiotic variables
```{r evp~abiotic}
# Summarize standardized/predicted evaporation data
# - Reminder: evaporation for each plot at 10am for every census date
d_evap = pred_dat %>%
  group_by(locID) %>%
  summarize(evap.1.avg = mean(pred.evap.1, na.rm = TRUE),
            evap.1.min = min(pred.evap.1),
            evap.1.max = max(pred.evap.1),
            evap.1.range = evap.1.max - evap.1.min)
# Load other measured independent variables
d_ind <- read.csv("data/roya_independentVariables.csv") %>%
  mutate(locID = as.factor(plotID))
# Join all abiotic variables together
d_allXs = d_evap %>%
  left_join(d_ind, by = "locID")

chart.Correlation(d_allXs %>%
                    select(evap.1.avg, evap.1.min, evap.1.max, evap.1.range, 
                           avgEvpRateConverted_g.mm2.s, # Compare methods
                           avgEvpRateConverted2_g.mm2.s, 
                           coffeeDensity_m, wind_m.s, shadeAvg_pc, 
                           avgTreeDist, tree_10m, tree_30m, tree_50m) %>%
                    # Removing outlier doesn't change correlations much
                    #filter(avgTreeDist < 15) %>% 
                    as.matrix(), histogram = TRUE, pch=19)
```

MH comments:

* All of the metrics calculated from standardized evaporation rate correlate strongly with each other (r = 1), for some reason. Why should a site with high average evap_std have higher evp range? (confirmed with plot below)
    * At least for now, I think we can consider these metrics interchangeable
* +wind +evap (r = 0.20 *), and the new metrics seem to match the original evaporation values well, better than my original conversion methods.
* We're still getting a weird, positive (significant) relationship between standardized evaporation metrics and tree density (r=0.26 **), especially at 30m radius. We expect a negative relationship, as +trees +shade -evaporation.
    * There is a weak positive relationship between shade and tree density

```{r graph - evap_std_avg~evap_std_range}
### Does increasing predicted evaporation mean increase range?
# Arrange histograms by increasing evap mean
pred_dat_2 = pred_dat %>%
  group_by(locID) %>%
  summarize(evap.1.avg = mean(pred.evap.1, na.rm = TRUE))

# Rank average evaporation rate and join with remaining data
rank_no = data.frame(rank_start = c(1:nrow(pred_dat_2))) %>%
  mutate(rank = ifelse(rank_start < 10, paste("00", rank_start, sep = ""),
                       ifelse(rank_start < 100, paste("0", rank_start, sep = ""),
                              rank_start)))

# Add ranking to dataset
pred_dat_2 = pred_dat_2 %>%
  arrange(evap.1.avg) %>%
  mutate(rank = rank_no$rank) %>%
  right_join(pred_dat, by = "locID") %>%
  mutate(rank = paste(rank, locID, sep="-"))

# Plot
ggplot(pred_dat_2) + 
  # geom_histogram(aes(x = pred.evap.0), color = "green", alpha = 0.2) + 
  geom_histogram(aes(x = pred.evap.1), color = "blue", alpha = 0.2) +
  facet_wrap(~rank) + theme_article()
```

MH: by ranking these histograms in order of increasing mean, we see that sites with higher mean standardized evaporation rate also has higher range of evaporation across days.

```{r evap~wind}
chart.Correlation(d_allXs %>%
                    select(evap.1.avg, evap.1.min, evap.1.max, evap.1.range, 
                           avgEvpRate_g.mm2.s, 
                           avgEvpRateConverted_g.mm2.s,
                           avgEvpRateConverted2_g.mm2.s,
                           wind_m.s) %>%
                    as.matrix(), histogram = TRUE, pch=19)

chart.Correlation(d_allXs %>%
                    select(evap.1.avg, evap.1.min, evap.1.max, evap.1.range, 
                           avgEvpRate_g.mm2.s, 
                           avgEvpRateConverted_g.mm2.s,
                           avgEvpRateConverted2_g.mm2.s,
                           wind_m.s) %>%
                    filter(wind_m.s > 0) %>%
                    as.matrix(), histogram = TRUE, pch=19)
```

MH: These plots show how the new standardized evaporation rates compare with my earlier attempts, by looking at their relationship with wind velocity. Note the relationship between wind and raw evaporation rate (avgEvpRate_g.mm2.s). When we remove wind = 0 values, we see that the new standardized evaporation rates have a similar correlation with wind (r = 0.35 **) as the raw data. My previous standardizing method (avgEvpRateConverted_g.mm2.s and avgEvpRateConverted2_g.mm2.s) do not match the original relationship as well.

# Look at the relationship between standardized evap and roya

```{r roya}
roya <- readRDS("./data/roya_uniqueplants_20210505.rds")

roya_quad_allXs6 <- roya %>% filter(hojas != 0) %>%
  group_by(quadrat, year, date) %>% 
  summarize(roya.mn = mean(roya/hojas, na.rm = TRUE)) %>% 
  filter(date > 2016.5 & date < 2017.5)

ggplot(roya_quad_allXs6, aes(date, roya.mn, group = quadrat)) + geom_line()

roya_quadmax <- roya_quad_allXs6 %>% group_by(quadrat) %>%
  summarize(roya.max = max(roya.mn)) %>%
  mutate(locID = as.character(quadrat))

evaproya <- evap_std %>% left_join(roya_quadmax)

ggplot(evaproya, aes(evap, roya.max)) + geom_point() + theme_article() + 
  geom_smooth(method = "lm")

evaplm <- lm(roya.max ~ evap, data = evaproya)

summary(evaplm)

saveRDS(evap_std, "evap_std.rds")

```